R for Clinical Study Reports and Submission

### 2021-10-06 project-management.Rmd
update project managment

SDLC for A&R deliverables can be defined in stages:

@@

The define stage is an important stage in the SDLC lifecycle 
as the requirements for all A&R deliverables are gathered and documented.

Importantly, we should not consider SDLC as a linear process. 
For example, if study team identify new requirement in a development or validation stage, 
team shall return to planning stage to discuss and align the scope. 
An [agile project management](https://www.atlassian.com/agile/project-management) approach is suitable 
and recommended to an A&R clinical development project.
The goal is to embrace an iterative approach that continue improve target deliverables based on
frequent customer feedback.

@@

@@

- mock-up
- validation level (e.g. independent review or double programming)
- etc.

ToDo: add a simplified mock-up and A&R grid.

@@

ToDo: add a simplified validation tracker

@@

To enable reproducibility, project leader shall also review the startup file 
(i.e. `.Rprofile` discussed in Section \@ref(manage)) and define:

@@

> After project initiate, modifying `.Rprofile` will be a risk of reproducibility 
> and shall be handled carefully if necessary.

@@

The analysis code and relevant description can be saved in R markdown files 
in the `vignettes` folder. 
The use of R markdown allow developer to assemble narrative text, code and its
comments in one place to simplify documentation. 
It would be helpful to define a name convention for all TLFs deliverables. 
For example, we can use `tlf_` prefix in the filename to indicate that the Rmarkdown is to deliver TLFs. 
For example, `tlf_001_ds.Rmd` is an Rmarkdown file for disposition table. 
Multiple TLFs with similar design can be included in one Rmarkdown file.

ToDo: add a file path for `esubdemo`

To validate the results, 
developer shall also save the last dataset before creating TLFs in RTF format to `.Rdata` files.

@@

@@

>>>>

SDLC for A&R deliverables can be defined in four stages:

@@

Importantly, we should not consider SDLC as a linear process. 
For example, if study team identify new requirement in a development or validation stage, 
team shall return to planning stage to discuss and align the scope. 
An [agile project management](https://www.atlassian.com/agile/project-management) approach is suitable 
and recommended to an A&R clinical development project.
The goal is to embrace an iterative approach that continue improve target deliverables based on
frequent stakeholder feedback.

@@

The planning stage is an important stage in the SDLC lifecycle 
as the requirements for all A&R deliverables are gathered and documented.

@@

- mock-up tables
- validation level (e.g. independent review or double programming)
- etc.

@@

```{r, out.width = "100%", out.height = "400px", echo = FALSE, fig.align = "center"}
knitr::include_graphics("images/validation_tracker.png")
```

@@

To enable reproducibility, project leader shall also review the startup file 
(i.e. `.Rprofile` discussed in Section \@ref(reproduce)) and define:

@@

> After project initiation, modifying `.Rprofile` will be a risk of reproducibility 
> and shall be handled carefully if necessary.

@@

The analysis code and relevant description can be saved in R markdown files 
in the `vignettes` folder.

The use of R markdown allow developer to assemble narrative text, code and its
comments in one place to simplify documentation. 
It would be helpful to create a template and define a name convention for all TLFs deliverables. 
For example, we can use `tlf_` prefix in the filename to indicate that the Rmarkdown is to deliver TLFs. 
Multiple TLFs with similar design can be included in one Rmarkdown file.

For example, in `esubdemo` project, we have 6 Rmarkdown files to create TLFs.

If there is any project specific R function that needs to be developed, 
the R functions can be placed in the `R\` folder as discussed in Section \@ref(consistency).

@@

Validation is an important stage to ensure the deliverables are accurate and consistent. After the
development stage is completed, the project team needs to validate the deliverables including Rmarkdown
files for TLFs deliverables and project specific R functions. The level of validation is determined at the define
stage.

In an R package development, the validation or testing is completed under the `test/` folder. 
The testthat R package can be used to streamline the validation process. 
More details of `testthat` package for R package validation can be found in Chapter 12 of the R package book.

It is recommended to have a name convention to indicate the type of validation. 
For example, we can use `test-developer-test`, `test-independent-test`, `test-double-programming` 
to classify the validation type.

It is recommended to follow the same organization
for files in testthat folder as `R/` folder and `vignettes/` folder. 
Every single file in the `R/` folder and `vignettes/` folder should have a testing file saved in 
the `tests/testthat/` folder to validate the content.

For example, in `esubdemo` project, we can have a list of testing files below.

```
tests/testthat
├── test-independent-test-tlf-01-disposition.Rmd
├── test-independent-test-tlf-02-population.Rmd
├── test-independent-test-tlf-03-baseline.Rmd
├── test-independent-test-tlf-04-efficacy.Rmd
├── test-independent-test-tlf-05-ae-summary.Rmd
├── test-independent-test-tlf-06-ae-spec.Rmd
└── test-independent-test-fmt.R
```

To validate the content of a Table, we can save the last datasets ready for table generation as a `.Rdata` file.
A validator can reproduces the TLF and compares it with the original result saved in `.Rdata` file. 
A test is passed when the results match. Customers can directly review the formatting of the TLFs by
customers by comparing with the mock-up.

To validate a Figure, we can use [snapshot testing](https://testthat.r-lib.org/articles/snapshotting.html) strategy

After the validator completes the testing of project specific functions and Rmarkdown files, 
the process to execute and report testing results is the same for a standard R package. 
The `devtools::test()` function automatically executes all testing cases and summarizes the testing results in a report.

After completing the validation, the validator updates the status in a validation tracker. 
The project lead reviews the tracking sheet to make sure all required activities in the SDLC are completed and
the tracking sheet has been filled correctly. The deliverables are ready for customer review after all the
validation steps are completed. Any changes to the output requested by customers are documented.

@@

After completion of development and required validation of all A&R deliverables, project lead runs compliance
check for a project specific R package similar to other R packages. Using `devtools::check()` is a
convenient way to run compliance check or R CMD checks. R CMD check is an automated check of the
contents in the R package for frequently encountered issues, before submission to CRAN. Since project
specific R package is not submitted to CRAN, some checks can be customized and skipped in
`devtools::check()`. The project lead shall work with study team to ensure all reported errors, warnings and
notes by `devtools::check()` are fixed.

The project lead can also use `pkgdown` R package to build a complete website for a project specific R
package. The `pkgdown` website is a convenient way to run all analysis in batch and 
integrate outputs in a website.

Project specific R functions, TLF generation programs, outputs and validation tracking information can all
be explored from the website.

For example, in `esubdemo` project, we created the `pkgdown` website in https://elong0527.github.io/esubdemo/.

Many of the tasks in SDLC can be completed automatically. 
An organization can leverage CI/CD workflow to automatically enable those tasks 
such as running testing cases and create a `pkgdown` website.

For example, in `esubdemo` project, we setup [github actions](https://github.com/elong0527/esubdemo/actions)
for these work. It can be setup by using `usethis::use_github_action()` function.
